Categories of Optics

### 2018-07-16 paper.tex
Fill in coalgebra section

\newcommand{\id}{\mathrm{id}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\const}{\mathrm{const}}
\DeclareMathOperator{\ob}{ob}
\DeclareMathOperator{\copr}{copr}
\newcommand{\inl}{\mathrm{inl}}
\newcommand{\inr}{\mathrm{inr}}
\DeclareMathOperator{\im}{im}

@@

If we ask more of our category $\C$, we can show that a lens $S \hto A$ implies the existence of a complement $\C$ with $S \cong C \times A$. This doesn't appear to follow purely from the concrete lens laws---an argument that a definition of lawfulness based on constant complements is not the correct generalisation.

@@

\begin{proposition}
  The category of optics for this action is isomorphic to $\C \times \C^\op$.
\end{proposition}
\begin{proof}
  \begin{align*}
    \Optic_1((S, S'), (A, A')) &= \int^{M \in 1} \C(S, MA) \times \C(MA', S') \\
                               &\cong \C(S, \star \cdot A) \times \C(\star \cdot A', S') \\
                               &\cong \C(S, A) \times \C(A', S')
  \end{align*}
  where $\star$ denotes the object of $1$. Composition in $\Optic_1((S, S'), (A, A'))$ does indeed correspond to composition in $\C \times \C^\op$.
\end{proof}

@@

There is a common pattern in some of the examples to follow: the evaluation-at-$A$ functor $-A : \M \to \C$ has a right adjoint, say $R_A : \C \to \M$.

In this case we can always eliminate the coend and find a concrete description of an optic:
\begin{align*}
&\int^{F \in \M} \C(S, F\cdot A) \times \C(F\cdot A', S') \\
&\cong \int^{F \in \M} \C(S, F \cdot A) \times \M(F, R_{A'} S') \\
&\cong \C(S,  (R_{A'} S') \cdot A)
\end{align*}

\begin{theorem}\label{thm:optics-are-coalgebras}
An concrete optic $p : S \to  (R_{A} S) \cdot A$ is lawful iff it is a coalgebra for the comonad $X \mapsto (R_{A} X) \cdot A$.
\end{theorem}
\begin{proof}
\todo{todo!}
\end{proof}

\begin{remark}
This theorem is a generalisation of \cite[Section 3.2]{RelatingAlgebraicAndCoalgebraic} for lenses.
\end{remark}

%Ross Street:
%More usually - . b has a right adjoint. This is the same as saying that the M-actegory
%C is an M-category (hom enriched in M) which is tensored. So we have
%
%M(m, \hat{C}(b,a)) \cong C(m.b , a)

@@

\begin{definition}[\cite{StrongFunctors}]\label{def:strong-functor}
  A \emph{(left) strong functor} is a functor $F : \C \to \C$ equipped with a natural transformation called the \emph{strength}:
  \begin{align*}
    \theta_{A,B} : A \otimes F B \to F(A \otimes B)
  \end{align*}
  such that the strength commutes with the unitor:
  \[
    \begin{tikzcd}
      I \otimes F A \ar[r, "\theta_{1,A}"] \ar[d, "\cong" left]  & F(I \otimes A) \ar[d, "\cong" right] \\
      F A \ar[r, equals] & F A
    \end{tikzcd}
  \]
  and with associativity:
  \[
    \begin{tikzcd}
      (A \otimes B) \otimes F C \ar[rr, "\theta_{A \otimes B, C}"] \ar[d, "\alpha_{A,B,FC}" left]  && F((A \otimes B) \otimes C) \ar[d, "F\alpha_{A,B,FC}" right] \\
      A \otimes (B \otimes F C) \ar[r, "A \otimes \theta_{B,C}" below] & A \otimes F(B \otimes C) \ar[r, "\theta_{A, B\otimes C}" below] & F(A \otimes (B \otimes C))
    \end{tikzcd}
  \]
  A \emph{strong natural transformation} $\tau : (F,\theta) \to (G,\theta')$ is a natural transformation that respects the strengths. There is an evident category $\Strong(\C)$ of strong endofunctors and strong natural transformations, and a forgetful functor $U : \Strong(\C) \to [\C, \C]$.
%  \[
%    \begin{tikzcd}
%      A \otimes F B \ar[r, "\theta_{A,B}"] \ar[d, "A \otimes \tau_B" left]  & F(A \otimes B) \ar[d, "\tau_{A \otimes B}" right] \\
%      A \otimes G B \ar[r, "\theta'_{A, B}"]& G(A \otimes B)
%    \end{tikzcd}
%  \]
%
\end{definition}
Then, again, $\Strong_\C$ acts on $\C$ by evaluation. We leave it to the reader to verify there is a natural isomorphism \[\C(S, FA) \cong \Strong_\C(\homC(A, -) \times S, F),\] which we can use to describe optics for this action as elements of $\C(\homC(A, A'), \homC(S,S'))$. The use of strong functors here is due to the correspondence between tensorial strengths and ``$\C$-enrichments'' on a functor.
\end{remark}

@@

\subsection{Traversals}
In this section we work in the case $\C = \Set$. Traversals allow us to traverse through a data structure, accumulating applicative (or monadic) actions as we go. We begin by reviewing the defnitions of applicative and traversable functors~\cite{AnInvestigationOfTheLawsOfTraversals}.

\begin{definition}
An \emph{applicative functor} $F : \C \to \C$ is a lax monoidal functor with a strength compatible with the monoidal structure, in the sense that
\[
\begin{tikzcd}[column sep = large]
A \otimes FB \otimes FC \ar[r, "{\theta_{A, B} \otimes FC}"] \ar[d, swap, "{A \otimes \phi_{B, C}}"] & F(A \otimes B) \otimes FC \ar[d, "{\phi_{A \otimes B, C}}"] \\
A \otimes F(B \otimes C) \ar[r, swap, "{\theta_{A, B \otimes C}}"] & F(A \otimes B \otimes C)
\end{tikzcd}
\]
commutes. An \emph{applicative natural transformation} is one that is both monoidal and strong. Applicative functors and natural transformations form a monoidal category $\App$ under functor composition.
\end{definition}

@@

We can then use the coalgebraic pattern from earlier to reach the same concrete description of traversals as found in~\cite{ProfunctorOptics}.
\begin{align*}
\Optic_\Traversable((S, S'), (A, A')) &= \int^{F \in \Traversable} \Set(S, F A) \times \Set(F A', S') \\
&\cong \int^{F \in \Traversable} \Set(S, F A) \times \Traversable(F, \Sigma_n {(-)}^n \times \Set(A'^n,S')) \\
&\cong \Set(S, \Sigma_n A^n \times \Set(A'^n,S'))
\end{align*}

@@

\begin{remark}
Traversable functors can be described as a particular class of polynomial functors known as finitary containers. It may be possible to generalise this to other classes of polynomial functors.
\end{remark}

@@

Polymorphic optics can be captured by the coend formalism as follows. Any action of a monoidal category $\M \times \C \to \C$ can be extended to act object-wise on a functor category:
\begin{align*}
  \M \times [\D, \C] &\to [\D, \C] \\
  (M, F) &\mapsto  M \cdot (F-)
\end{align*}

@@

If $\C$ is closed monoidal, but not necessarily cartesian, we can use the tensor-hom adjunction to eliminate the coend:

\begin{align*}
  \Optic_\otimes((S, S'), (A, A')) &= \int^{M \in \C} \C(S, M \otimes A) \times \C(M \otimes A', S') \\
                                   &\cong \int^{M \in \C} \C(S, M \otimes A) \times \C(M, \homC(A',S')) \\
                                   &\cong \C(S, \homC(A',S') \otimes A)
\end{align*}
where $\homC(A', S')$ denotes the internal hom. If $\C$ is cartesian, this is of course isomorphic to the pair of maps discussed earlier.

A concrete linear lens is therefore a map $\funzip : S \to \homC(A',S') \otimes A$. The above isomorphism sends $\funzip$ to the element $\rep{\funzip}{\ev_{A', S'}}$, where $\ev_{A', S'} : \homC(A',S') \otimes A' \to S'$ is the evaluation map. In the other direction, given $\rep{l}{r}$ we have $\funzip = (\tilde r \otimes A)l$ where $\tilde r : M \to \homC(A', S')$ denotes the transpose of $r$.

We cannot possibly use the three $\fput$/$\fget$ style lens laws in this setting as we lack projections, but using the same techniques as for cartesian lenses above we can devise some new ones:

\todo{These are just the coalgebra laws (hopefully)}

\begin{proposition}\label{prop:concrete-linear-lawful}
  A linear lens $p : S \hto A$ is lawful iff the following two concrete laws for $\funzip$ hold:
  \begin{align*}
    \ev_{A, S} \; \funzip &= \id_S && \textsc{(Rezip)} \\
    (\coev_{\homC(A, S), A} \otimes A)\funzip &= ((\funzip \circ -) \otimes A)\funzip && \textsc{(ZipZip)}
  \end{align*}
  where \[ \funzip \circ - : \homC(A, S) \to \homC(A, \homC(A, S) \otimes A) \] denotes internal composition and \[\coev_{\homC(A, S), A} : \homC(A, S) \to \homC(A, \homC(A, S) \otimes A)\] is coevaluation.
  \todo{I obviously need better names for the laws}
\end{proposition}
\begin{proof}
  Let \[\overrightarrow{(-)} : \C(A \otimes B, C) \to \C(A, \homC(B, C)) : \overleftarrow{(-)}\] denote the homset bijection, so 
  \begin{align*}
    \coev_{A, B} &:= \overrightarrow{\id_{A \otimes B}} \\
    \ev_{A, B} &:= \overleftarrow{\id_{\homC(A, B)}}
  \end{align*}

  We have an isomorphism:
  \begin{align*}
  \Twoptic_\otimes(S, A)
    &= \int^{M_1, M_2 \in \C} \C(S, M_1 \otimes A) \times \C(M_1 \otimes A, M_2 \otimes A) \times \C(M_2 \otimes A, S) \\
    &\cong \int^{M_1, M_2 \in \C} \C(S, M_1 \otimes A) \times \C(M_1, \homC(A, M_2 \otimes A)) \times \C(M_2, \homC(A, S)) \\
    &\cong \int^{M_1 \in \C} \C(S, M_1 \otimes A) \times \C(M_1, \homC(A, \homC(A, S) \otimes A)) \\
    &\cong \C(S, \homC(A, \homC(A, S) \otimes A) \otimes A)
  \end{align*}
  which evaluated on an element $\repthree{l}{c}{r}$ is given by
  \begin{align*}
    \Phi(\repthree{l}{c}{r}) &= (\overrightarrow{(\overrightarrow{r} \otimes A)c} \otimes A)l
  \end{align*}
  Now suppose we are given an optic $p$ whose corresponding element of $\conc_\otimes(S, A)$ is $\funzip$. Then \[\outside(\rep{\funzip}{\ev_{A, S} }) = \ev_{A, S} \; \funzip,\] giving the first law.

  Now we evaluate $\Phi(\once(p))$ and $\Phi(\twice(p))$.
  \begin{align*}
    \Phi(\once(\rep{\funzip}{\ev_{A, S}}))
    &= \Phi(\repthree{\funzip}{\id_{\homC(A,S) \otimes A}}{\ev_{A, S} }) \\
    &= (\overrightarrow{(\overrightarrow{\ev_{A, S}} \otimes A)\id_{\homC(A,S) \otimes A}} \otimes A)\funzip \\
    &= (\overrightarrow{(\id_{\homC(A, S)} \otimes A)} \otimes A)\funzip \\
    &= (\coev_{\homC(A, S), A} \otimes A)\funzip \\
    \Phi(\twice(\rep{\funzip}{\ev_{A, S} }))
    &= \Phi(\repthree{\funzip}{\funzip \; \ev_{A, S}}{\ev_{A, S} }) \\
    &= (\overrightarrow{(\overrightarrow{\ev_{A, S}} \otimes A)(\funzip \; \ev_{A, S})} \otimes A)\funzip \\
    &= (\overrightarrow{(\funzip \; \ev_{A, S})} \otimes A)\funzip \\
    &= ((\funzip \circ -) \otimes A)\funzip
  \end{align*}
  So indeed the optic is lawful iff the two concrete laws hold.
\end{proof}

@@

\begin{proposition}
If $T : \C \to \C$ is a strong monad then $\C$ acts on $\C_T$ by $X \cdot Y := X \otimes Y$.
\end{proposition}

@@

\begin{itemize}
\item \emph{``Achromatic'' Lenses}~\cite[Section 5.2]{ProfunctorOpticsThesis} are lenses that also admit an operation $\fcreate : A \to S$. These are optics for the action of $\C$ on itself by $M \cdot A = (M \sqcup 1) \times A$, or equivalently, of the category of pointed objects of $\C$ on $\C$ by cartesian product. Concrete achromatic lenses $S \hto A$ are elements of the set \[\C(S, \homC(A', S') \sqcup 1) \times \C(S, A) \times \C(A', S').\]
%\begin{align*}
%  \Optic((S, S'), (A, A'))
%  &= \int^{M \in \C} \C(S, M \cdot A) \times \C(M \cdot A', S') \\
%  &= \int^{M \in \C} \C(S, (M \sqcup 1) \times A) \times \C((M \sqcup 1) \times A', S') \\
%  &\cong \int^{M \in \C} \C(S, (M \sqcup 1) \times A) \times \C((M \times A') \sqcup A', S') \\
%  &\cong \int^{M \in \C} \C(S, (M \sqcup 1) \times A) \times \C(M \times A', S') \times \C(A', S') \\
%  &\cong \int^{M \in \C} \C(S, (M \sqcup 1) \times A) \times \C(M, \homC(A', S')) \times \C(A', S') \\
%  &\cong \C(S, (\homC(A', S') \sqcup 1) \times A) \times \C(A', S') \\
%  &\cong \C(S, \homC(A', S') \sqcup 1) \times \C(S, A) \times \C(A', S')
%\end{align*}
\item \emph{Affine Traversals}~\cite{AffineTraversalPost} allow access to a target that may or may not be present. Suppose $\C$ is cartesian closed and has binary coproducts. Let $\mathsf{Aff}$ be the category $\C \times \C$, equipped with the monoidal structure
\begin{align*}
  (P', Q') \otimes (P, Q) &= (P' \sqcup (Q' \times P) , Q' \times Q)
\end{align*}
The category $\mathsf{Aff}$ acts on $\C$ by $(P, Q) \cdot A = P \sqcup (Q \times A)$, in fact, $\mathsf{Aff}$ is cooked up exactly as the closure of the actions $- \times A$ and $- \sqcup A$ under composition. A concrete affine traversal is an element of \[\C(S, S' \sqcup (\homC(A', S') \times A)).\]

@@

%\begin{align*}
%  (P', Q') \cdot (P, Q) \cdot A
%  &= (P', Q') \cdot  (P \sqcup (Q \times A)) \\
%  &= P' \sqcup (Q' \times (P \sqcup (Q \times A))) \\
%  &\cong P' \sqcup (Q' \times P) \sqcup (Q' \times Q \times A) \\
%  &= (P' \sqcup (Q' \times P) , Q' \times Q) \cdot A \\
%  &= ((P', Q') \otimes (P, Q)) \cdot A
%\end{align*}
%\begin{remark}
%  It is important here that the morphisms in $\mathsf{Aff}$ are only those that arise from pairs of morphisms $P \to P'$ and $Q \to Q'$, although in principle there may be other natural transformations between the corresponding functors $(P, Q) \cdot -$ and $(P', Q') \cdot -$.
%\end{remark}
%\todo{This feels similar to taking some sort of `compositum' of the two actions $\times$ and $\sqcup$, both embed in this category. Marco suggests taking the pushout of the projections into the pullback, calculated in the 2-cat of monoidal categories.}
%Now the set of optics $(S, S') \hto (A, A')$ is:
%\begin{align*}
%  \Optic_{\mathsf{Aff}}((S, S'), (A, A'))
%  &= \int^{M \in \mathsf{Aff}} \C(S, M \cdot A) \times \C(M \cdot A', S') \\
%  &\cong \int^{P,Q \in \C} \C(S, (P,Q) \cdot A) \times \C((P,Q) \cdot A', S') \\
%  &= \int^{P,Q \in \C} \C(S, P \sqcup (Q \times A)) \times \C(P \sqcup (Q \times A'), S') \\
%  &\cong \int^{P,Q \in \C} \C(S, P \sqcup (Q \times A)) \times \C(P,S') \times \C(Q \times A', S') \\
%  &\cong \int^{Q \in \C} \C(S, S' \sqcup (Q \times A)) \times \C(Q \times A', S') \\
%  &\cong \int^{Q \in \C} \C(S, S' \sqcup (Q \times A)) \times \C(Q, \homC(A', S')) \\
%  &\cong \C(S, S' \sqcup (\homC(A', S') \times A))
%\end{align*}

\item \emph{Adjunction Lenses}: For a category with finite products $\C$ and a monoidal adjunction $F \dashv U$, with $F : \C \to \D$ and $U : \D \to \C$. If $U$ is strong monoidal then $\D$ acts on $\C$ by $UD \cdot A := D \times A$; as an example consider the free-forgetful adjunction between sets and monoids. Concrete optics for this action are given pairs in
\[ \C(S, A) \times \C(UFS \times A', S') \]
\todo{This example is a bit dumb, delete?}
\end{itemize}

@@

\begin{proposition}
The assignment $F \cdot X \mapsto F \Set(I, X)$ is a lax monoidal action of $\App$ on $\Set$.
\end{proposition}
Call this the $I$-indexed action of $\App$ on $\Set$.

@@

\begin{proposition}
$\IdxTraversable$ acts on $\C \times \C$ by $(T, I) \cdot (A, R) \mapsto (TA, \homC(I, R))$ and on $\C$ simply by $(T, I) \cdot A \mapsto TA$.
\end{proposition}

>>>>

\newcommand{\id}{\mathrm{id}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\const}{\mathrm{const}}
\DeclareMathOperator{\ob}{ob}
\DeclareMathOperator{\copr}{copr}
\newcommand{\inl}{\mathrm{inl}}
\newcommand{\inr}{\mathrm{inr}}
\DeclareMathOperator{\im}{im}
\newcommand{\act}{\cdot}

@@

If we ask more of our category $\C$, we can show that a lens $S \hto A$ implies the existence of a complement $\C$ with $S \cong C \times A$.  This doesn't appear to follow purely from the concrete lens laws---an argument that a definition of lawfulness based on constant complements is not the correct generalisation. For completeness we include a proof in our notation.

@@

\begin{proposition}
  The category of optics for this action is isomorphic to $\C \times \C^\op$.
\end{proposition}
\begin{proof}
  \begin{align*}
    \Optic_1((S, S'), (A, A')) &= \int^{M \in 1} \C(S, MA) \times \C(MA', S') \\
                               &\cong \C(S, \star \act A) \times \C(\star \act A', S') \\
                               &\cong \C(S, A) \times \C(A', S')
  \end{align*}
  where $\star$ denotes the object of $1$. Composition in $\Optic_1((S, S'), (A, A'))$ does indeed correspond to composition in $\C \times \C^\op$.
\end{proof}

@@

There is a common pattern in many of the examples to follow: for every object $A \in \C$, the evaluation-at-$A$ functor $-A : \M \to \C$ has a right adjoint, say $R_A : \C \to \M$. To fix notation, let \[\overrightarrow{(-)} : \C(F \act A, S) \to \M(F, R_{A} S) : \overleftarrow{(-)}\] denote the homset bijection, so the unit and counit are:
  \begin{align*}
  	\eta_F &: F \to R_A (F \act A) \\
    \eta_F &:= \overrightarrow{\id_{F \act A}} \\
    \varepsilon_{S} &: (R_A S) \cdot A \to S \\
    \varepsilon_{S} &:= \overleftarrow{\id_{R_{A} S}}
  \end{align*}

\begin{remark}
It is shown in~\cite[Section 6]{ANoteOnActions} that, at least in the case $\M$ is right closed, to give such an action is equivalent to giving a ``copowered $\M$-category $\C$''. \todo{Pointed out to me years ago when I emailed Ross Street, but I didn't understand what he was saying}
\end{remark}

In this case we can always eliminate the coend and find a concrete description of an optic:
\begin{align*}
&\int^{F \in \M} \C(S, F\act A) \times \C(F\act A', S') \\
&\cong \int^{F \in \M} \C(S, F \act A) \times \M(F, R_{A'} S') \\
&\cong \C(S,  (R_{A'} S') \act A)
\end{align*}
A concrete optic is therefore a map $\funzip : S \to (R_{A'} S') \act A$. The above isomorphism sends $\funzip$ to the element $\rep{\funzip}{\varepsilon_{S'}}$. In the other direction, given $\rep{l}{r}$ we have $\funzip = (\overrightarrow{r} \otimes A)l$.

\begin{theorem}\label{thm:optics-are-coalgebras}
An concrete optic $\funzip : S \to  (R_{A} S) \act A$ is lawful iff it is a coalgebra for the comonad $X \mapsto (R_{A} X) \act A$.
\end{theorem}
\begin{proof}
By adjointness and Yoneda reduction we have an isomorphism $\Phi : \Twoptic_\M(S, A) \to \C(S, R_A (R_A S \act A) \act A)$ given by
\begin{align*}
&\Twoptic_\M(S, A) \\
&= \int^{M_1, M_2 \in \M} \C(S, M_1 \act A) \times \C(M_1 \act A, M_2 \act A) \times \C(M_2 \act A, S) \\
&\cong \int^{M_1, M_2 \in \M} \C(S, M_1 \act A) \times \M(M_1, R_A (M_2 \act A)) \times \M(M_2, R_A S) \\
&\cong \int^{M_1, M_2 \in \M} \C(S, M_1 \act A) \times \M(M_1, R_A (M_2 \act A)) \times \M(M_2, R_A S) \\
&\cong \int^{M_1 \in \M} \C(S, M_1 \act A) \times \M(M_1, R_A (R_A S \act A)) \\
&\cong \C(S, R_A (R_A S \act A) \act A)
\end{align*}
  which evaluated on an element $\repthree{l}{c}{r}$ is given by
  \begin{align*}
    \Phi(\repthree{l}{c}{r}) &= (\overrightarrow{(\overrightarrow{r} \act A)c} \act A)l
  \end{align*}

So now interpreting the optic laws, find
\[\outside(\rep{\funzip}{\varepsilon_{S} }) = \varepsilon_{S} \; \funzip = \id_S \] is exactly the coalgebra unit law, and equality of
  \begin{align*}
    \Phi(\once(\rep{\funzip}{\varepsilon_S}))
    &= \Phi(\repthree{\funzip}{\id_{(R_{A} S) \act A}}{\varepsilon_S }) \\
    &= (\overrightarrow{(\overrightarrow{\varepsilon_S} \act A)\id_{(R_{A} S) \act A}} \act A)\funzip \\
    &= (\overrightarrow{(\id_{R_A S} \act A)} \act A)\funzip \\
    &= (\overrightarrow{\id_{R_A S \act A}} \act A)\funzip \\
    &= (\eta_{R_A S} \act A) \funzip \\
    \Phi(\twice(\rep{\funzip}{\varepsilon_S }))
    &= \Phi(\repthree{\funzip}{\funzip \; \varepsilon_S}{\varepsilon_S }) \\
    &= (\overrightarrow{(\overrightarrow{\varepsilon_S} \act A)(\funzip \; \varepsilon_S)} \act A)\funzip \\
    &= (\overrightarrow{(\funzip \; \varepsilon_S)} \act A)\funzip \\
    &= (R_A (\funzip) \act A)\funzip
  \end{align*}
is exactly the coalgebra multiplication law.
\end{proof}

@@

\begin{definition}[\cite{StrongFunctors}]\label{def:strong-functor}
  A \emph{(left) strong functor} is a functor $F : \C \to \C$ equipped with a natural transformation called the \emph{strength}:
  \begin{align*}
    \theta_{A,B} : A \otimes F B \to F(A \otimes B)
  \end{align*}
  such that the strength commutes with the unitor:
  \[
    \begin{tikzcd}
      I \otimes F A \ar[r, "\theta_{1,A}"] \ar[d, "\cong" left]  & F(I \otimes A) \ar[d, "\cong" right] \\
      F A \ar[r, equals] & F A
    \end{tikzcd}
  \]
  and with associativity:
  \[
    \begin{tikzcd}
      (A \otimes B) \otimes F C \ar[rr, "\theta_{A \otimes B, C}"] \ar[d, "\alpha_{A,B,FC}" left]  && F((A \otimes B) \otimes C) \ar[d, "F\alpha_{A,B,FC}" right] \\
      A \otimes (B \otimes F C) \ar[r, "A \otimes \theta_{B,C}" below] & A \otimes F(B \otimes C) \ar[r, "\theta_{A, B\otimes C}" below] & F(A \otimes (B \otimes C))
    \end{tikzcd}
  \]
  A \emph{strong natural transformation} $\tau : (F,\theta) \to (G,\theta')$ is a natural transformation that respects the strengths. There is an evident category $\Strong(\C)$ of strong endofunctors and strong natural transformations, and a forgetful functor $U : \Strong(\C) \to [\C, \C]$.
%  \[
%    \begin{tikzcd}
%      A \otimes F B \ar[r, "\theta_{A,B}"] \ar[d, "A \otimes \tau_B" left]  & F(A \otimes B) \ar[d, "\tau_{A \otimes B}" right] \\
%      A \otimes G B \ar[r, "\theta'_{A, B}"]& G(A \otimes B)
%    \end{tikzcd}
%  \]
%
\end{definition}
Then, again, $\Strong_\C$ acts on $\C$ by evaluation. We leave it to the reader to verify there is a natural isomorphism \[\C(S, FA) \cong \Strong_\C(\homC(A, -) \times S, F),\] which we can use to describe optics for this action as elements of $\C(\homC(A, A'), \homC(S,S'))$. 
%The use of strong functors here is due to the correspondence between tensorial strengths and ``$\C$-enrichments'' on a functor.
\end{remark}

@@

\subsection{Traversals}
In this section we work in the case $\C = \Set$. Traversals allow us to traverse through a data structure, accumulating applicative actions as we go. We begin by reviewing the defnitions of applicative and traversable functors~\cite{AnInvestigationOfTheLawsOfTraversals}.

\begin{definition}
An \emph{applicative functor} $F : \C \to \C$ is a lax monoidal functor with a strength compatible with the monoidal structure, in the sense that
\[
\begin{tikzcd}[column sep = large]
A \otimes FB \otimes FC \ar[r, "{\theta_{A, B} \otimes FC}"] \ar[d, swap, "{A \otimes \phi_{B, C}}"] & F(A \otimes B) \otimes FC \ar[d, "{\phi_{A \otimes B, C}}"] \\
A \otimes F(B \otimes C) \ar[r, swap, "{\theta_{A, B \otimes C}}"] & F(A \otimes B \otimes C)
\end{tikzcd}
\]
commutes. An \emph{applicative natural transformation} is one that is both monoidal and strong. Applicative functors and natural transformations form a monoidal category $\App$ with the tensor given by functor composition.
\end{definition}

@@

We can then use the coalgebraic pattern from earlier to reach the same concrete description of traversals as found in~\cite{ProfunctorOptics}.
\begin{align*}
\Optic_\Traversable((S, S'), (A, A')) &\cong \Set(S, \Sigma_n A^n \times \Set(A'^n,S'))
\end{align*}

@@

%\begin{remark}
%Traversable functors can be described as a particular class of polynomial functors known as finitary containers. It may be possible to generalise this to other classes of polynomial functors.
%\end{remark}

@@

Polymorphic optics can be captured by the coend formalism as follows. Any action of a monoidal category $\M \times \C \to \C$ can be extended to act object-wise on a functor category:
\begin{align*}
  \M \times [\D, \C] &\to [\D, \C] \\
  (M, F) &\mapsto  M \act (F-)
\end{align*}

@@

If $\C$ is closed monoidal but not necessarily cartesian, we can still define the category of \emph{linear lenses} to be $\Optic_\otimes$. The internal hom provides a right adjoint to the evaluation-at-$A$ functor, so we have immediately
\begin{align*}
  \Optic_\otimes((S, S'), (A, A')) &\cong \C(S, \homC(A',S') \otimes A)
\end{align*}
where $\homC(A', S')$ denotes the internal hom. If $\C$ is cartesian, this is of course isomorphic to the set of $(\fget, \fput)$ functions discussed earlier.

We cannot possibly use the three $\fput$/$\fget$ style lens laws in this setting as we lack projections, but specialising the coalgebra laws gives us:

\begin{proposition}\label{prop:concrete-linear-lawful}
  A linear lens $p : S \hto A$ is lawful iff the following two concrete laws for $\funzip$ hold:
  \begin{align*}
    \ev_{A, S} \; \funzip &= \id_S && \textsc{(Rezip)} \\
    (\coev_{\homC(A, S), A} \otimes A)\funzip &= ((\funzip \circ -) \otimes A)\funzip && \textsc{(ZipZip)}
  \end{align*}
  where \[ \funzip \circ - : \homC(A, S) \to \homC(A, \homC(A, S) \otimes A) \] denotes internal composition and \[\coev_{\homC(A, S), A} : \homC(A, S) \to \homC(A, \homC(A, S) \otimes A)\] is coevaluation.
  \todo{I obviously need better names for the laws} \qed
\end{proposition}

\begin{remark}
We have essentially rederived the result given in \cite[Section 3.2]{RelatingAlgebraicAndCoalgebraic} for ordinary lenses, but we note that cartesianness was not required.
\end{remark}

@@

\begin{proposition}
If $T : \C \to \C$ is a strong monad then $\C$ acts on $\C_T$ by $X \act Y := X \otimes Y$.
\end{proposition}

@@

\begin{itemize}
\item \emph{``Achromatic'' Lenses}~\cite[Section 5.2]{ProfunctorOpticsThesis} are lenses that also admit an operation $\fcreate : A \to S$. These are optics for the action of $\C$ on itself by $M \act A = (M \sqcup 1) \times A$, or equivalently, of the category of pointed objects of $\C$ on $\C$ by cartesian product. Concrete achromatic lenses $S \hto A$ are elements of the set \[\C(S, \homC(A', S') \sqcup 1) \times \C(S, A) \times \C(A', S').\]
%\begin{align*}
%  \Optic((S, S'), (A, A'))
%  &= \int^{M \in \C} \C(S, M \act A) \times \C(M \act A', S') \\
%  &= \int^{M \in \C} \C(S, (M \sqcup 1) \times A) \times \C((M \sqcup 1) \times A', S') \\
%  &\cong \int^{M \in \C} \C(S, (M \sqcup 1) \times A) \times \C((M \times A') \sqcup A', S') \\
%  &\cong \int^{M \in \C} \C(S, (M \sqcup 1) \times A) \times \C(M \times A', S') \times \C(A', S') \\
%  &\cong \int^{M \in \C} \C(S, (M \sqcup 1) \times A) \times \C(M, \homC(A', S')) \times \C(A', S') \\
%  &\cong \C(S, (\homC(A', S') \sqcup 1) \times A) \times \C(A', S') \\
%  &\cong \C(S, \homC(A', S') \sqcup 1) \times \C(S, A) \times \C(A', S')
%\end{align*}
\item \emph{Affine Traversals}~\cite{AffineTraversalPost} allow access to a target that may or may not be present. Suppose $\C$ is cartesian closed and has binary coproducts. Let $\mathsf{Aff}$ be the category $\C \times \C$, equipped with the monoidal structure
\begin{align*}
  (P', Q') \otimes (P, Q) &= (P' \sqcup (Q' \times P) , Q' \times Q)
\end{align*}
The category $\mathsf{Aff}$ acts on $\C$ by $(P, Q) \act A = P \sqcup (Q \times A)$, in fact, $\mathsf{Aff}$ is cooked up exactly as the closure of the actions $- \times A$ and $- \sqcup A$ under composition. A concrete affine traversal is an element of \[\C(S, S' \sqcup (\homC(A', S') \times A)).\]

@@

%\begin{align*}
%  (P', Q') \act (P, Q) \act A
%  &= (P', Q') \act  (P \sqcup (Q \times A)) \\
%  &= P' \sqcup (Q' \times (P \sqcup (Q \times A))) \\
%  &\cong P' \sqcup (Q' \times P) \sqcup (Q' \times Q \times A) \\
%  &= (P' \sqcup (Q' \times P) , Q' \times Q) \act A \\
%  &= ((P', Q') \otimes (P, Q)) \act A
%\end{align*}
%\begin{remark}
%  It is important here that the morphisms in $\mathsf{Aff}$ are only those that arise from pairs of morphisms $P \to P'$ and $Q \to Q'$, although in principle there may be other natural transformations between the corresponding functors $(P, Q) \act -$ and $(P', Q') \act -$.
%\end{remark}
%\todo{This feels similar to taking some sort of `compositum' of the two actions $\times$ and $\sqcup$, both embed in this category. Marco suggests taking the pushout of the projections into the pullback, calculated in the 2-cat of monoidal categories.}
%Now the set of optics $(S, S') \hto (A, A')$ is:
%\begin{align*}
%  \Optic_{\mathsf{Aff}}((S, S'), (A, A'))
%  &= \int^{M \in \mathsf{Aff}} \C(S, M \act A) \times \C(M \act A', S') \\
%  &\cong \int^{P,Q \in \C} \C(S, (P,Q) \act A) \times \C((P,Q) \act A', S') \\
%  &= \int^{P,Q \in \C} \C(S, P \sqcup (Q \times A)) \times \C(P \sqcup (Q \times A'), S') \\
%  &\cong \int^{P,Q \in \C} \C(S, P \sqcup (Q \times A)) \times \C(P,S') \times \C(Q \times A', S') \\
%  &\cong \int^{Q \in \C} \C(S, S' \sqcup (Q \times A)) \times \C(Q \times A', S') \\
%  &\cong \int^{Q \in \C} \C(S, S' \sqcup (Q \times A)) \times \C(Q, \homC(A', S')) \\
%  &\cong \C(S, S' \sqcup (\homC(A', S') \times A))
%\end{align*}

\item \emph{Adjunction Lenses}: For a category with finite products $\C$ and a monoidal adjunction $F \dashv U$, with $F : \C \to \D$ and $U : \D \to \C$. If $U$ is strong monoidal then $\D$ acts on $\C$ by $UD \act A := D \times A$; as an example consider the free-forgetful adjunction between sets and monoids. Concrete optics for this action are given pairs in
\[ \C(S, A) \times \C(UFS \times A', S') \]
\todo{This example is a bit dumb, delete?}
\end{itemize}

@@

\begin{proposition}
The assignment $F \act X \mapsto F \Set(I, X)$ is a lax monoidal action of $\App$ on $\Set$.
\end{proposition}
Call this the $I$-indexed action of $\App$ on $\Set$.

@@

\begin{proposition}
$\IdxTraversable$ acts on $\C \times \C$ by $(T, I) \act (A, R) \mapsto (TA, \homC(I, R))$ and on $\C$ simply by $(T, I) \act A \mapsto TA$.
\end{proposition}
